Empirical Analysis of Performance Bottlenecks in Graph Neural Network Training with GPUs

### 2020-12-24 NEUROC/Response-MajRev/response_letter.tex

\begin{reply}
    Thank you for your positive comments on our manuscript.
    %
    We have carefully revised the manuscript according to your comments. 
    %
    Based on the suggestions, we have made an extensive modification on the revised manuscript.
    %
    Please see the detailed responses below.
    %
    The changes corresponding to each issue were highlighted by red squares in the annotated version of our manuscript.

\end{reply}

% R1Q1
\begin{point}
	There are lots of symbols in this paper. Some symbols are reused and confusing, such as s denotes sub-layers or edge features.
\end{point}

% Our reply
\begin{reply}
    We apologize for the confusing use of symbols and thank you for pointing the problem out.
    %
    To clarify the symbol usage, we have checked the manuscript and unified the usage of symbol.
    %
    We avoid the problem of reusing symbols, so that each symbol only represents one meaning after the revision.
    %
    We summarize the frequently-used symbols in Table 1 in the revised manuscript.
    %
    We quote the table below:
    \begin{table}[H]
        \footnotesize
        \centering
        \begin{tabular}{p{3em}lp{35em}}
            \toprule
            Category & Symbol & Meaning \\
            \midrule
            \multirow[c]{4}{3em}{Graph Structure}& $\mathcal{G}=(\mathcal{V}, \mathcal{E})$ & The simple undirected input graph with the vertex set $\mathcal{V}$ and the edge set $\mathcal{E}$. \\
            & $v_x$ & The $x$-th vertex of the input graph. \\
            & $e_{y,x}$ & The edge pointing from $v_y$ to $v_x$ of the input graph. \\
            & $\mathcal{N}(v_x)$ & The adjacency set of $v_x$ in the input graph. \\ 
            & $\bar{d}$ & The average degree of the input graph. \\ \midrule
            \multirow[c]{6}{3em}{GNN Definition}& $L$ & The number of GNN layers. \\
            & $K$ & The number of heads in a GNN layer. \\
            & $\phi^l$ & The messaging function of the GNN layer $l$. \\
            & $\Sigma^l$ & The aggregation function of the GNN layer $l$. \\
            & $\gamma^l$ & The vertex updating function of the GNN layer $l$. \\ 
            & $\phi^{l,i}$ / $\Sigma^{l,i}$ / $\gamma^{l,i}$ & The messaging/aggregation/updating function of the $i$-th sub-layer of the GNN layer  $l$.\\
            & $\textcolor{blue}{\boldsymbol{W}^l, \boldsymbol{W}^{(k)}/\boldsymbol{b}, \boldsymbol{a}}$ & The matrices/vectors represented by the blue characters are the weight matrices/vectors that need to be learned in the GNN. \\  \midrule
            \multirow[c]{8}{3em}{Vector}& $\boldsymbol{v}_x$ & The feature vector of the vertex $v_x$. \\
            & $\boldsymbol{e}_{y,x}$ & The feature vector of the edge $e_{y,x}$.  \\
            & $\boldsymbol{h}_x^{l}$ &  The {input} hidden vector of the graph neuron corresponding to $v_x$ in the GNN layer $l$. \\
            & $\boldsymbol{h}_x^{l+1}$ &  The {output} hidden vector of the graph neuron corresponding to $v_x$ in the GNN layer $l$.\\
            & $\boldsymbol{m}_{y,x}^l$ & The message vector of the edge $e_{y,x}$ outputted by $\phi^l$ of the GNN layer $l$. \\
            & $\boldsymbol{s}_{x}^l$ & The aggregated vector of the vertex $v_x$ outputted by $\Sigma^l$ of the GNN layer $l$. \\
            & $\boldsymbol{h}_{x}^{l,i}$ / $\boldsymbol{m}_{y,x}^{l,i}$ / $\boldsymbol{s}_{x}^{l,i}$ & The hidden/message/aggregated vector of the vertex $v_x$ outputted by $\gamma^{l,i}$/$\phi^{l,i}$/$\Sigma^{l,i}$ of the $i$-th sub-layer of the GNN layer $l$. \\
            & $d^l_{in}$, $d^l_{out}$ &  The dimension of the input/output hidden vectors of the GNN layer $l$. \\
            & $dim(\MyVec{x})$ & The dimension of a vector $\MyVec{x}$. \\
            \bottomrule
        \end{tabular}
\end{table}

    In the revised manuscript, we use $e_{y,x}$ to represent an edge and use $\boldsymbol{e}_{y,x}$ to represent its input feature vector.
    %
    The input feature vectors of all edges are same for all GNN layers.
    %
    We use $\boldsymbol{s}$ to represent aggregated vectors outputted by the aggregation function $\Sigma$ in graph neurons.
    %
    For every vertex $v_x$, we use $\boldsymbol{s}^l_x$ to denote its aggregated vector in the GNN layer $l$.
    %
    If the GNN layer $l$ has sub-layers, $\boldsymbol{s}^{l,i}_x$ represents its aggregated vector in the $i$-th sub-layer.

### 2020-12-24 NEUROC/sec2-review-of-graph-neural-network.tex
将e_{x,y}表示与后文统一

In this section, we formally define the graph neural networks and briefly survey typical graph neural networks.
%
We denote a simple \emph{undirected} graph $\mathcal{G}$ as $\mathcal{G}=(\mathcal{V}, \mathcal{E})$, where $\mathcal{V}$ and $\mathcal{E}$ are the vertex set and the edge set of $\mathcal{G}$, respectively.
%
We use $v_x$ $(0 \leq x < |\mathcal{V}|)$ to denote a vertex and $e_{x,y}=(v_x, v_y)$ to denote the edge pointing from $v_x$ to $v_y$.
%
The adjacency set of $v_x$ is $\mathcal{N}(v_x)=\{v|(v_x, v) \in \mathcal{E}\}$.
%
We denote a \emph{vector} with a bold lower case letter like $\boldsymbol{h}$ and a \emph{matri}x with a bold upper case letter like $\boldsymbol{W}$.
%
Table~\ref{tab:notations} summarizes the common symbols used throughout this work.
%
We use blue characters $\Param{\MyVec{w}}$/$\Param{\MyMat{W}}$ to denote weight vectors/matrices that are model parameters to train in a GNN.

\begin{table}[h]
    \caption{Notations}
    \label{tab:notations}
    \centering
    \small
    \begin{tabular}{p{3em}lp{35em}}
        \toprule
       Category & Symbol & Meaning \\
        \midrule
        \multirow[c]{4}{3em}{Graph Structure}& $\mathcal{G}=(\mathcal{V}, \mathcal{E})$ & The simple undirected input graph with the vertex set $\mathcal{V}$ and the edge set $\mathcal{E}$. \\
        & $v_x$ & The $x$-th vertex of the input graph. \\
        & $e_{x,y}$ & The edge pointing from $v_x$ to $v_y$ of the input graph. \\
        & $\mathcal{N}(v_x)$ & The adjacency set of $v_x$ in the input graph. \\ 
        & $\bar{d}$ & The average degree of the input graph. \\ \midrule
        \multirow[c]{6}{3em}{GNN Definition}& $L$ & The number of GNN layers. \\
        & $K$ & The number of heads in a GNN layer. \\
        & $\phi^l$ & The messaging function of the GNN layer $l$. \\
        & $\Sigma^l$ & The aggregation function of the GNN layer $l$. \\
        & $\gamma^l$ & The vertex updating function of the GNN layer $l$. \\ 
        & $\phi^{l,i}$ / $\Sigma^{l,i}$ / $\gamma^{l,i}$ & The messaging/aggregation/updating function of the $i$-th sub-layer of the GNN layer  $l$.\\
        & $\textcolor{blue}{\boldsymbol{W}^l, \boldsymbol{W}^{(k)}/\boldsymbol{b}, \boldsymbol{a}}$ & The matrices/vectors represented by the blue characters are the weight matrices/vectors that need to be learned in the GNN. \\  \midrule
        \multirow[c]{8}{3em}{Vector}& $\boldsymbol{v}_x$ & The feature vector of the vertex $v_x$. \\
        & $\boldsymbol{e}_{x,y}$ & The feature vector of the edge $e_{x,y}$.  \\
        & $\boldsymbol{h}_x^{l}$ &  The {input} hidden vector of the graph neuron corresponding to $v_x$ in the GNN layer $l$. \\
        & $\boldsymbol{h}_x^{l+1}$ &  The {output} hidden vector of the graph neuron corresponding to $v_x$ in the GNN layer $l$.\\
        & $\boldsymbol{m}_{x,y}^l$ & The message vector of the edge $e_{x,y}$ outputted by $\phi^l$ of the GNN layer $l$. \\
        & $\boldsymbol{s}_{x}^l$ & The aggregated vector of the vertex $v_x$ outputted by $\Sigma^l$ of the GNN layer $l$. \\
        & $\boldsymbol{h}_{x}^{l,i}$ / $\boldsymbol{m}_{x,y}^{l,i}$ / $\boldsymbol{s}_{x}^{l,i}$ & The hidden/message/aggregated vector of the vertex $v_x$ outputted by $\gamma^{l,i}$/$\phi^{l,i}$/$\Sigma^{l,i}$ of the $i$-th sub-layer of the GNN layer $l$. \\
        & $d^l_{in}$, $d^l_{out}$ &  The dimension of the input/output hidden vectors of the GNN layer $l$. \\
        & $dim(\MyVec{x})$ & The dimension of a vector $\MyVec{x}$. \\
        \bottomrule
    \end{tabular}
\end{table}

@@

In the inpug layer, A GNN receives a graph $\mathcal{G}$ as the input.
%
Every vertex $v_x$ in $\mathcal{G}$ is attached with a feature vector $\boldsymbol{v}_x$ to describe the properties of the vertex.
%
Every edge $e_{x,y}$ of $\mathcal{G}$ may also be attached with a feature vector $\boldsymbol{e}_{x,y}$.
%
The input layer of a GNN receives feature vectors from all vertices and passes the feature vectors to the first GNN layer (i.e. GNN layer 0).

A GNN usually consists of more than one GNN layers.
%
Each GNN layer consists of $|\mathcal{V}|$ graph neurons, where $|\mathcal{V}|$ is the number of vertices in $\mathcal{G}$.
%
Each graph neuron corresponds to a vertex in $\mathcal{G}$.
%
Different GNNs mainly differ in the graph neurons that they use.
%
We elaborate on details of graph neurons later.
%
GNN layers are \emph{sparsely} connected with the input layer and other GNN layers.
%
\begin{itemize}
    \item In the first GNN layer (layer 0), a graph neuron collects feature vectors from the input layer.
    %
    For the graph neuron corresponding to the vertex $v_x$, it collects the feature vector $\boldsymbol{v}_x$ and the feature vectors $\boldsymbol{v}_y$ of the vertices $v_y$ that are adjacent to $v_x$ (i.e.  $v_y \in \mathcal{N}(v_x)$).
    %
    The graph neuron aggregates input feature vectors, apply non-linear transformation, and outputs a hidden vector $\boldsymbol{h}^1_x$ for $v_x$.
    %
    Take the demo GNN in \figurename~\ref{fig:general_structure_of_gnn}(a) as an example.
    %
    Since $\mathcal{N}(v_3) = \{v_1, v_2, v_4, v_5, v_6\}$, the graph neuron of $v_3$ at the GNN layer 0 collects the input feature vectors \{$\boldsymbol{v}_1$, $\boldsymbol{v}_2$, $\boldsymbol{v}_3$, $\boldsymbol{v}_4$, $\boldsymbol{v}_5$, $\boldsymbol{v}_6$\} from the input layer and outputs $\boldsymbol{h}^1_3$.
    %
    The connections between the first GNN layer and the input layer  are determined by the topology of $\mathcal{G}$.
    %
    The graph neuron of $v_x$ in the first GNN layer is connected with the input unit of $v_y$  in the input layer only if there is an edge $e_{x,y}$ between $v_x$ and $v_y$ in $\mathcal{G}$.
    %
    Since most real-world graphs are very \emph{sparse} (i.e. $|\mathcal{E}| \ll |\mathcal{V}|^2$), the connections between the first GNN layer and the input layer are also sparse, different from the traditional neural networks.

>>>>

In this section, we formally define the graph neural networks and briefly survey typical graph neural networks.
%
We denote a simple \emph{undirected} graph $\mathcal{G}$ as $\mathcal{G}=(\mathcal{V}, \mathcal{E})$, where $\mathcal{V}$ and $\mathcal{E}$ are the vertex set and the edge set of $\mathcal{G}$, respectively.
%
We use $v_x$ $(0 \leq x < |\mathcal{V}|)$ to denote a vertex and $e_{y,x}=(v_y, v_x)$ to denote the edge pointing from $v_y$ to $v_x$.
%
The adjacency set of $v_x$ is $\mathcal{N}(v_x)=\{v|(v, v_x) \in \mathcal{E}\}$.
%
We denote a \emph{vector} with a bold lower case letter like $\boldsymbol{h}$ and a \emph{matri}x with a bold upper case letter like $\boldsymbol{W}$.
%
Table~\ref{tab:notations} summarizes the common symbols used throughout this work.
%
We use blue characters $\Param{\MyVec{w}}$/$\Param{\MyMat{W}}$ to denote weight vectors/matrices that are model parameters to train in a GNN.

\begin{table}[h]
    \caption{Notations}
    \label{tab:notations}
    \centering
    \small
    \begin{tabular}{p{3em}lp{35em}}
        \toprule
       Category & Symbol & Meaning \\
        \midrule
        \multirow[c]{4}{3em}{Graph Structure}& $\mathcal{G}=(\mathcal{V}, \mathcal{E})$ & The simple undirected input graph with the vertex set $\mathcal{V}$ and the edge set $\mathcal{E}$. \\
        & $v_x$ & The $x$-th vertex of the input graph. \\
        & $e_{y,x}$ & The edge pointing from $v_y$ to $v_x$ of the input graph. \\
        & $\mathcal{N}(v_x)$ & The adjacency set of $v_x$ in the input graph. \\ 
        & $\bar{d}$ & The average degree of the input graph. \\ \midrule
        \multirow[c]{6}{3em}{GNN Definition}& $L$ & The number of GNN layers. \\
        & $K$ & The number of heads in a GNN layer. \\
        & $\phi^l$ & The messaging function of the GNN layer $l$. \\
        & $\Sigma^l$ & The aggregation function of the GNN layer $l$. \\
        & $\gamma^l$ & The vertex updating function of the GNN layer $l$. \\ 
        & $\phi^{l,i}$ / $\Sigma^{l,i}$ / $\gamma^{l,i}$ & The messaging/aggregation/updating function of the $i$-th sub-layer of the GNN layer  $l$.\\
        & $\textcolor{blue}{\boldsymbol{W}^l, \boldsymbol{W}^{(k)}/\boldsymbol{b}, \boldsymbol{a}}$ & The matrices/vectors represented by the blue characters are the weight matrices/vectors that need to be learned in the GNN. \\  \midrule
        \multirow[c]{8}{3em}{Vector}& $\boldsymbol{v}_x$ & The feature vector of the vertex $v_x$. \\
        & $\boldsymbol{e}_{y,x}$ & The feature vector of the edge $e_{y,x}$.  \\
        & $\boldsymbol{h}_x^{l}$ &  The {input} hidden vector of the graph neuron corresponding to $v_x$ in the GNN layer $l$. \\
        & $\boldsymbol{h}_x^{l+1}$ &  The {output} hidden vector of the graph neuron corresponding to $v_x$ in the GNN layer $l$.\\
        & $\boldsymbol{m}_{y,x}^l$ & The message vector of the edge $e_{y,x}$ outputted by $\phi^l$ of the GNN layer $l$. \\
        & $\boldsymbol{s}_{x}^l$ & The aggregated vector of the vertex $v_x$ outputted by $\Sigma^l$ of the GNN layer $l$. \\
        & $\boldsymbol{h}_{x}^{l,i}$ / $\boldsymbol{m}_{y,x}^{l,i}$ / $\boldsymbol{s}_{x}^{l,i}$ & The hidden/message/aggregated vector of the vertex $v_x$ outputted by $\gamma^{l,i}$/$\phi^{l,i}$/$\Sigma^{l,i}$ of the $i$-th sub-layer of the GNN layer $l$. \\
        & $d^l_{in}$, $d^l_{out}$ &  The dimension of the input/output hidden vectors of the GNN layer $l$. \\
        & $dim(\MyVec{x})$ & The dimension of a vector $\MyVec{x}$. \\
        \bottomrule
    \end{tabular}
\end{table}

@@

In the inpug layer, A GNN receives a graph $\mathcal{G}$ as the input.
%
Every vertex $v_x$ in $\mathcal{G}$ is attached with a feature vector $\boldsymbol{v}_x$ to describe the properties of the vertex.
%
Every edge $e_{y,x}$ of $\mathcal{G}$ may also be attached with a feature vector $\boldsymbol{e}_{y,x}$.
%
The input layer of a GNN receives feature vectors from all vertices and passes the feature vectors to the first GNN layer (i.e. GNN layer 0).

A GNN usually consists of more than one GNN layers.
%
Each GNN layer consists of $|\mathcal{V}|$ graph neurons, where $|\mathcal{V}|$ is the number of vertices in $\mathcal{G}$.
%
Each graph neuron corresponds to a vertex in $\mathcal{G}$.
%
Different GNNs mainly differ in the graph neurons that they use.
%
We elaborate on details of graph neurons later.
%
GNN layers are \emph{sparsely} connected with the input layer and other GNN layers.
%
\begin{itemize}
    \item In the first GNN layer (layer 0), a graph neuron collects feature vectors from the input layer.
    %
    For the graph neuron corresponding to the vertex $v_x$, it collects the feature vector $\boldsymbol{v}_x$ and the feature vectors $\boldsymbol{v}_y$ of the vertices $v_y$ that are adjacent to $v_x$ (i.e.  $v_y \in \mathcal{N}(v_x)$).
    %
    The graph neuron aggregates input feature vectors, apply non-linear transformation, and outputs a hidden vector $\boldsymbol{h}^1_x$ for $v_x$.
    %
    Take the demo GNN in \figurename~\ref{fig:general_structure_of_gnn}(a) as an example.
    %
    Since $\mathcal{N}(v_3) = \{v_1, v_2, v_4, v_5, v_6\}$, the graph neuron of $v_3$ at the GNN layer 0 collects the input feature vectors \{$\boldsymbol{v}_1$, $\boldsymbol{v}_2$, $\boldsymbol{v}_3$, $\boldsymbol{v}_4$, $\boldsymbol{v}_5$, $\boldsymbol{v}_6$\} from the input layer and outputs $\boldsymbol{h}^1_3$.
    %
    The connections between the first GNN layer and the input layer  are determined by the topology of $\mathcal{G}$.
    %
    The graph neuron of $v_x$ in the first GNN layer is connected with the input unit of $v_y$  in the input layer only if there is an edge $e_{y,x}$ between $v_y$ and $v_x$ in $\mathcal{G}$.
    %
    Since most real-world graphs are very \emph{sparse} (i.e. $|\mathcal{E}| \ll |\mathcal{V}|^2$), the connections between the first GNN layer and the input layer are also sparse, different from the traditional neural networks.
